Enforced Invariant Confluence

### 2018-04-01 doc/paper/sections/segmented_invariant_confluence.tex
Finished segmented invariant-confluence section.

% why does this work

\subsection{Discussion}
% generalizes locking
% contiuous generalization of invariant-confluence

% The motivate the main idea behind segmented invariant-confluence, let's revisit
% \exampleref{Z2} but with a different start state.
%
% \begin{example}
%   Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
%   = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
%   \exampleref{Z2}, but now let our start state $s_0$ be $(-10, 10)$. Now, $O$
%   is \emph{not} \sTIconfluent{}. From $(-10, 10)$, we can reach the points $(0,
%   10)$ and $(10, 0)$, and merging these together we reach $(10, 10)$ which
%   violates the invariant.
%   %
%   However, all is not lost. If we're determined to make $O$
%   invariant-confluent, we can do so in one of two ways.
%   \begin{enumerate}
%     \item
%       We can modify the invariant. $O$ may not be \sTIconfluent, but it is
%       \sticonfluent{s_0}{T}{\setst{(x, y)}{x \leq 0, y \geq 0}}. If we eliminate
%       the fourth quadrant from the invariant, $O$ becomes invariant-confluent.
%     \item
%       We can disallow some transactions. $O$ may not be \sTIconfluent, but
%       it is \sticonfluent{s_0}{\set{t_{x+1}}}{I}. If we disallow the $t_{y-1}$
%       transaction, $O$ becomes invariant-confluent.
%   \end{enumerate}
%   While these two modifications are possible, neither is desirable. Artificially making some invariant-satisfying states unreachable---either by restricting the invariant or disallowing some transactions---is most likely a bit of a hack. Plus,
% \end{example}

>>>>

This system model guarantees that all replicas of a segmented
invariant-confluent object always satisfy the invariant. All servers begin in
the same initial state and with the same active segment. Thus, because $O$ is
invariant-confluent with respect to every segment, servers can execute
transactions within the active segment without any coordination and guarantee
that the invariant is never violated. Any operation that would violate the
assumptions of invariant-confluence system model (e.g.\ executing a transaction
that's not permitted in the active segment or executing a permitted transaction
that leads to a state outside the active segment) triggers a global
coordination. Globally coordinated transactions are only executed if they
maintain the invariant. Moreover, if a globally coordinated transaction leads
to another segment, the coordination ensures that all servers begin in the same
start state and with the same active replica, reestablishing the assumptions of
the invariant-confluence system model.

\subsection{Discussion}
There are a couple of things worth noting about segmented invariant-confluence
and its system model. First, invariant-confluence is a very coarse-grained
property. If an object is invariant-confluent, then we can replicate it with no
coordination. If it is not invariant-confluent, then we have no guarantees.
There's no in-between. Segmented invariant-confluence, on the other hand, is a
much more fine-grained property that can be applied to applications with
varying degrees of complexity.  Segmented invariant-confluence provides
guarantees to complex applications that require a large number of segments and
to simple applications with a smaller number of segments, whereas
invariant-confluence only provides guarantees to applications that can be
segmented into a single segment.

Second, segmented invariant-confluence can naturally model distributed locking.
One approach to replicating non-invariant-confluent objects is to recognize
which transactions cannot be safely executed concurrently and then require them
to acquire a distributed lock before executing~\cite{balegas2015putting,
gotsman2016cause}. For example, in a banking application with the invariant
that all balances remain positive, concurrent deposits are permitted, but
concurrent withdrawals can lead to invariant violations, thus forcing
withdrawals to first acquire a distributed lock before executing. To model this
locking with segmented invariant-confluence, we simply remove lock-acquiring
transactions from every segment's set of transactions. Consequently, executing
these transactions are executed with global coordination, achieving the same
effect as acquiring a distributed lock. In fact, segmented invariant-confluence
can be used to model more fine-grained locking. Omitting a transaction from
only certain segments is analogous to requiring a transaction to acquire a
distributed lock only when certain conditions are met.

% TODO: Forward reference examples in the evaluation section.

Third, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination required. First, if a server with
state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
to execute, and $t(s_i)$ violates the active invariant but not the global
invariant, instead of initiating a round of global coordination, $p_i$ can
simply buffer $t$ for re-execution at a later time. While this increases the
latency required to execute $t$, it's possible that after other transactions
are executed, re-executing $t$ may lead to a state that either satisfies the
active invariant or violates the global invariant. In either case, a round of
global coordination is avoided. Similarly, servers can buffer transactions that
require global coordination, executing an entire batch of these transactions
during a single round of global coordination.
